Analyzing Factors Affecting Pakistan Cricket Team's Performance (1952-2024)

### 2024-12-11 paper/paper.qmd
Updated typos and added missing reference

---
title: "Analyzing Factors Affecting Pakistan Cricket Team's Performance (1952-2024)"
subtitle: "Opponent and Coin Tossplay Play Key Role in Deciding Outcome"
author: 
  - Muhammad Abdullah Motasim
thanks: "Code and data are available at: [https://github.com/abdullah-motasim/Analyzing-Pakistans-Cricket-Data](https://github.com/abdullah-motasim/Analyzing-Pakistans-Cricket-Data)."
date: today
date-format: long
abstract: "This paper analyzes Pakistan's cricket performance from 1952 to 2024, using a linear regression model to examine factors influencing match outcomes. We find that opponent strength and toss outcomes are significant predictors of Pakistan's success, highlighting both internal team dynamics and external variables. These findings provide information for coaches, analysts, and administrators, offering strategies to improve performance under varying conditions. Ultimately, our analysis contributes to a deeper understanding of the factors that drive Pakistan’s cricket success, with implications for teams worldwide."
format: pdf
number-sections: true
bibliography: references.bib
table-of-contents: true
toc-depth: 2
editor: 
  markdown: 
    wrap: 72
---

@@

library(tidyverse)
library(here)
library(ggplot2)
library(arrow)
library(modelsummary)
library(tinytable)
library(knitr)
library(kableExtra)

@@

Pakistan’s cricket team holds a storied legacy, marked by remarkable highs and challenging lows. From their 1992 World Cup triumph to fluctuations in form over the decades, analyzing trends in performance offers information into their journey and areas for strategic improvement. Despite the wealth of player statistics and match highlights available, there has been a lack of systematic statistical analysis that evaluates long-term performance trends. This paper fills that gap by analyzing data from 2002 to 2024 to uncover the key factors that contribute to Pakistan’s success in international cricket.

We employ a linear regression model to analyze match outcomes,
considering variables like year, opponent, and toss outcomes. Our findings reveal several important observations. First, the strength of the opposing team emerged as a significant factor influencing Pakistan’s chances of winning. We also found that toss outcomes had a noteworthy impact on match results. This paper highlights the importance of both internal factors (such as team performance) and external factors (such as opponent strength) in shaping match outcomes. These observations provide strategic guidance for coaches, analysts, and cricket administrators, not only in Pakistan but also around the world, on how to prepare and perform under varying conditions.

The structure of this paper is as follows: @sec-data discusses the data
types included in the raw data, the cleaning process for the data, and
the reason for selecting the data set we did. @sec-model discusses model
specification and justification for a Linear Regression model.
@sec-results analyzes the trends and correlations between different
variables utilizing tabular and graphical means. @sec-discussion
discusses the results of @sec-results going into detail on what the
simulation results can tell us about Pakistan cricket performance, as
well as assumptions and limitations in data.

@@

The data for this study were obtained from the cricketdata R package
[@cricketdata] and all analysis was performed using `R` [@citeR]
alongside the following packages: `tibble` [@tibble], `readr` [@readr],
`arrow` [@arrow], `tidyverse` [@tidyverse], `dplyr` [@dplyr], `testthat`
[@testthat], `caret` [@caret], `rstanarm` [@rstanarm], `tidyr` [@tidyr],
modelsummary [@modelsummary], tinytable [@tinytable], knitr [@knitr], and ggplot2 [@ggplot2].

@@

Figures @fig-match-stats show key match statistics for Pakistan. The data shows that Pakistan frequently faces England, Sri Lanka, and Australia, with a particularly poor record against Australia, including several draws.

@@

-   `Prior for the Auxiliary Parameter (prior_aux)`: This represents the dispersion of the logistic regression model and is assigned an Exponential prior with a rate of 1, typical for Bayesian logistic regression. It reflects uncertainty in outcome variance, with higher values suggesting more variability.

@@

Our results are summarized in @tbl-model-results, and @fig-model-results
presents the predicted values of the coefficients. This figure also
includes the 95% confidence intervals for the predictors, as estimated
by our Bayesian model. A 95% confidence interval indicates that there is
a 95% probability that the true value of the parameter lies within the
specified range, given the observed data. In the graph, each point
represents the predicted value, and the horizontal line extending from
each point shows the 95% confidence interval.

@@

# table1 <- fit_diagnostics |>
#   tt()|>
#   style_tt(j = 2:5, align = "r") |>   # Align columns 2 to 5 to the right
#   format_tt(digits = 1, num_fmt = "decimal")

@@

# fit_diagnostics2 |>
#   tt() |>
#   style_tt(j = 2:4, align = "r") |>
#   format_tt(digits = 1, num_fmt = "decimal")

>>>>

---
title: "Analyzing Factors Affecting Pakistan Cricket Team's Performance (1952-2024)"
subtitle: "Opponent and Coin Toss Play Key Role in Deciding Match Outcome"
author: 
  - Muhammad Abdullah Motasim
thanks: "Code and data are available at: [https://github.com/abdullah-motasim/Analyzing-Pakistans-Cricket-Data](https://github.com/abdullah-motasim/Analyzing-Pakistans-Cricket-Data)."
date: today
date-format: long
abstract: "This paper analyzes Pakistan's cricket performance from 1952 to 2024, using a Bayesian Logistic Regression model to examine factors influencing match outcomes. We find that opponent strength and toss outcomes are significant predictors of Pakistan's success. These findings provide information for coaches, analysts, and administrators, offering strategies to improve performance under varying conditions. Ultimately, our analysis contributes to a deeper understanding of the factors that drive Pakistan’s cricket success, with implications for teams worldwide."
format: pdf
number-sections: true
bibliography: references.bib
table-of-contents: true
toc-depth: 2
editor: 
  markdown: 
    wrap: 72
---

@@

library(tidyverse)
library(here)
library(ggplot2)
library(arrow)
library(modelsummary)
library(knitr)
library(kableExtra)

@@

Pakistan’s cricket team holds a storied legacy, marked by remarkable highs and challenging lows. From their 1992 World Cup triumph to fluctuations in form over the decades, analyzing trends in performance offers information into their journey and areas for strategic improvement. Despite the wealth of player statistics and match highlights available, there has been a lack of systematic statistical analysis that evaluates long-term performance trends. This paper fills that gap by analyzing data from 1952 to 2024 to uncover the key factors that contribute to Pakistan’s success in international cricket.

We employ a Bayesian Logistic Regression model to analyze match outcomes,
considering variables like year, opponent, and toss outcomes. Our findings reveal several important observations. First, the strength of the opposing team emerged as a significant factor influencing Pakistan’s chances of winning. We also found that toss outcomes had a noteworthy impact on match results. This paper highlights the importance of both internal factors (such as team performance) and external factors (such as opponent strength) in shaping match outcomes. These observations provide strategic guidance for coaches, analysts, and cricket administrators, not only in Pakistan but also around the world, on how to prepare and perform under varying conditions.

The structure of this paper is as follows: @sec-data discusses the data
types included in the raw data, the cleaning process for the data, and
the reason for selecting the data set we did. @sec-model discusses model
specification and justification for a Bayesian Logistic Regression model.
@sec-results analyzes the trends and correlations between different
variables utilizing tabular and graphical means. @sec-discussion
discusses the results of @sec-results going into detail on what the
simulation results can tell us about Pakistan cricket performance, as
well as assumptions and limitations in data.

@@

The data for this study were obtained from the cricketdata R package
[@cricketdata] and all analysis was performed using `R` [@citeR]
alongside the following packages: `tibble` [@tibble], `readr` [@readr],
`arrow` [@arrow], `tidyverse` [@tidyverse], `dplyr` [@dplyr], `testthat`
[@testthat], `caret` [@caret], `rstanarm` [@rstanarm], `tidyr` [@tidyr],
`modelsummary` [@modelsummary], `knitr` [@knitr], `kableExtra` [@kableExtra] and `ggplot2` [@ggplot2].

@@

@fig-match-stats show key match statistics for Pakistan. The data shows that Pakistan frequently faces England, Sri Lanka, and Australia, with a particularly poor record against Australia, including several draws.

@@

-   `Prior for the Auxiliary Parameter (prior_aux)`: This represents the dispersion of the logistic regression model and is assigned an Exponential prior with a rate of 1, typical for Bayesian Logistic Regression. It reflects uncertainty in outcome variance, with higher values suggesting more variability.

@@

Our results are summarized in @tbl-model-results and @fig-model-results
presenting the predicted values of the coefficients. The figure also
includes the 95% confidence intervals for the predictors, as estimated
by our Bayesian model. A 95% confidence interval indicates that there is
a 95% probability that the true value of the parameter lies within the
specified range, given the observed data. In the graph, each point
represents the predicted value, and the horizontal line extending from
each point shows the 95% confidence interval.

@@

@@
